---
title: "final project"
output: github_document
date: "2024-10-02"
---
```{r}
library(dada2)
packageVersion("dada2")
```

## CHANGE ME to the directory containing the fastq files after unzipping.
```{r}
path <- "/home/rstudio/MiSeq_SOP" 
list.files(path)
```

# Forward and reverse fastq filenames have format: SAMPLENAME_R1_001.fastq and SAMPLENAME_R2_001.fastq
```{r}
fnFs <- sort(list.files(path, pattern="_R1_001.fastq", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_001.fastq", full.names = TRUE))
# Extract sample names, assuming filenames have format: SAMPLENAME_XXX.fastq
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
```

```{r}
plotQualityProfile(fnFs[1:2])
```

```{r}
plotQualityProfile(fnRs[1:2])
```


# Place filtered files in filtered/ subdirectory
```{r}
filtFs <- file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))
names(filtFs) <- sample.names
names(filtRs) <- sample.names
```

```{r}
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(240,160),
              maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
              compress=TRUE, multithread=TRUE) # On Windows set multithread=FALSE
head(out)
```


```{r}
errF <- learnErrors(filtFs, multithread=TRUE)
```

```{r}
errR <- learnErrors(filtRs, multithread=TRUE)
```

```{r}
plotErrors(errF, nominalQ=TRUE)
```

```{r}
dadaFs <- dada(filtFs, err=errF, multithread=TRUE)
```

```{r}
dadaRs <- dada(filtRs, err=errR, multithread=TRUE)
```

```{r}
dadaFs[[1]]
```

```{r}
mergers <- mergePairs(dadaFs, filtFs, dadaRs, filtRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
```
```{r}
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
```
```{r}
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
```

```{r}
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
```

```{r}
sum(seqtab.nochim)/sum(seqtab)
```

```{r}
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(dadaRs, getN), sapply(mergers, getN), rowSums(seqtab.nochim))
# If processing a single sample, remove the sapply calls: e.g. replace sapply(dadaFs, getN) with getN(dadaFs)
colnames(track) <- c("input", "filtered", "denoisedF", "denoisedR", "merged", "nonchim")
rownames(track) <- sample.names
head(track)
```
# Charger la base de données de taxonomie pour la classification
```{r}
taxa_file <- "path/to/silva_nr_v132_train_set.fa.gz"
```


# Vérifiez que le fichier existe
```{r}

if (file.exists(taxa_file)) {
    taxa <- assignTaxonomy(seqtab.nochim, taxa_file, multithread=TRUE)
    print("Taxonomy assigned successfully!")
} else {
    print("Taxonomy file not found!")
}
```


```{r}

```

```
```{r}
print(head(taxa))
```


```{r}
taxa.print <- taxa # Removing sequence rownames for display only
rownames(taxa.print) <- NULL
head(taxa.print)
```

